---
title: "EDA"
output: html_notebook
---
```{r include=FALSE}
library(forecast)
library(padr)
library(treemap)
library(lubridate)
```

# `holidays`
```{r}
summary(
  holidays %>% mutate(date = as.Date(date),
                      type = as.factor(type),
                      locale = as.factor(locale),
                      locale_name = as.factor(locale_name))
        )
```

```{r}
holidays %>% filter(type=="Work Day")
```

```{r}
holidays %>% group_by(date) %>% summarise(n=n()) %>%
  filter(n > 1)
```

No missing value in `holidays`
```{r}
colSums(is.na(holidays))
```

# `items`
```{r}
lapply(items, function(x) length(unique(x)))
```
No missing value in `items`
```{r}
colSums(is.na(items))
```
All classes are attached to unique family. We can consider family as a parent set of class.
```{r}
items %>% group_by(class) %>%
  summarise(in_family = length(unique(family))) %>%
  filter(in_family > 1)
```
It seems that all perishable item are having the class between 2000 and 3000
```{r}
ggplot(items) +
  geom_count(aes(y=family, x=class, color = perishable))
```
```{r}
paste("the class of perishable items are between ",
  min(items[items$perishable != 0,"class"]),
  " to ",
  max(items[items$perishable != 0,"class"])
)
```

`oil`
*43 missing value in oil price*
```{r}
colSums(is.na(oil))
```
combine with `holidays` to check
```{r}
oil %>% left_join(holidays, by="date") %>%
  filter(is.na(dcoilwtico))
```

PLot gold price over time
- a sharp fall in the second half of 2014
- a decresing trend from 2015 to 2016
- slowing increasing trend starting from 2016
```{r}
ggplot(pad(oil %>% mutate(date),interval = "day")) +
  geom_line(aes(x=date, y=dcoilwtico)) +
  ggtitle("Oil price between 2013-01-01 and 2017-08-31")
```

`stores`
no missing value
```{r}
colSums(is.na(stores))
```

- for citys, Quito (18) and Guayaquil (8) are the most among all
- for state, Pichincha(19) and Guayas(11) are the most among all
- only cluster 3,6,10 & 15 has 5 or more observations, the rest could be considered to combine
```{r}
lapply(stores[,c(2:5)],function(x) table(x))
```
More than 80% state has less than 3 stores
```{r}
stores %>% group_by(state) %>% summarise(n=n()) %>% ungroup %>%
  group_by(Store_nr_in_1_state=n) %>% summarise(count = n()) %>%
  mutate(perc = count/sum(count)*100) 
```
Only 4 states have stores in more than 1 city
```{r}
stores %>% group_by(state, city) %>% summarise(nr_store_city=n()) %>% ungroup() %>%
  group_by(state) %>% mutate(nr_store_state = sum(nr_store_city)) %>% 
  arrange(-nr_store_state, -nr_store_city) %>% filter(nr_store_city != nr_store_state)
```
Only cluster 10 has multiple types, we can consider cluster as a subset of type
```{r}
ggplot(stores) +
  geom_count(aes(x = as.factor(cluster), y=type, color = state)) +
  scale_x_discrete(name = "Cluster")
```

`train`
onpromotion has many missing value => check how onpromotion correlates to other factors
```{r}
colSums(is.na(train))
```
Number of items on sales by shop
```{r}
overall <- train %>% group_by(store_nbr, item_nbr) %>%
  summarise(total_sales = sum(unit_sales),
            item_ava = n()) %>%
  left_join(stores, by="store_nbr") %>%
  left_join(items, by="item_nbr")
```
Total unit sales by shop
```{r}
ggplot(overall)+
  geom_bar(aes(x=store_nbr, y=total_sales, fill = type), stat="identity") +
  scale_y_continuous(labels = function(x) paste(x/1000000,"M"), 
                     name = "Total unit sales") +
  scale_x_discrete(name = "Store") +
  ggtitle("Number of items on sales by store")
```
Obviously, shop with more item avaliable has higher unit sales
```{r}
ggplot(overall %>% group_by(store_nbr) %>% 
         summarise(total_sales = sum(total_sales),
                   item_ava = sum(item_ava))) +
  geom_point(aes(x = item_ava, y=total_sales)) +
  scale_y_continuous(labels = function(x) paste(x/1000000,"M"))+
  scale_x_continuous(labels = function(x) paste(x/1000000,"M"))
```

Total unit sales by state
```{r}
ggplot(overall)+
  geom_bar(aes(x=state, y=total_sales, fill = type), stat="identity") +
  scale_y_continuous(labels = function(x) paste(x/1000000,"M"), 
                     name = "Total unit sales") +
  scale_x_discrete(name = "State")+
  ggtitle("Number of items on sales by state")
```

Over sales unit by cluster
```{r}
ggplot(overall)+
  geom_bar(aes(x=as.factor(cluster), y=total_sales, fill = type), stat="identity") +
  scale_y_continuous(labels = function(x) paste(x/1000000,"M"), 
                     name = "Total unit sales") +
  scale_x_discrete(name = "Cluster") +
  ggtitle("Total unit sales by cluster")
```
Type A and D have the greatest items on sales
```{r}
ggplot(overall)+
  geom_bar(aes(x=type, y=total_sales), stat="identity") +
  scale_y_continuous(labels = function(x) paste(x/1000000,"M"), 
                     name = "Total unit sales") +
  scale_x_discrete(name = "type") +
  ggtitle("Number of items on sales by type")
```
Grocery I has most item to be sold. Beverages comes to the 2nd.
```{r}
ggplot(overall)+
  geom_bar(aes(x=family, y=total_sales), stat="identity") +
  scale_y_continuous(labels = function(x) paste(x/1000000,"M"), 
                     name = "Total unit sales") +
  scale_x_discrete(name = "family") +
  ggtitle("Number of items available by family") +
  coord_flip()
```

```{r}
train %>% group_by(year, month) %>% summarise(sales = sum(item_nbr))
str(train)
```
Top 20 items with most unit sales
```{r}
ggplot(head(train %>% group_by(item_nbr) %>%
         summarise(unit_sales = sum(unit_sales)) %>%
         arrange(-unit_sales),20) %>%
         left_join(items, by="item_nbr")) +
  geom_bar( aes(x=reorder(item_nbr, -unit_sales), y=unit_sales, fill = family), stat = "identity") +
  ggtitle("Top 20 unit sales items") + 
  scale_y_continuous(labels = function(x) paste(x/1000000,"M"), name = "unit sales") +
  scale_x_discrete(name = "Item Nr") +
  coord_flip()
  
```
transaction
```{r}
colSums(is.na(trans))
```
Total transaction
- the total number of transcations has an increasing trend
- the seasonality in 2016 is slight different to other 
```{r}
ggplot(trans %>% group_by(month, year) %>% 
         summarise(trans = sum(transactions)) %>% 
         ungroup) +
  geom_line(aes(x=as.numeric(month), y=trans, color = as.factor(year))) +
  ggtitle("Total number of transactions") +
  scale_y_continuous(name = "Transactions") +
  scale_x_continuous(name = "Month",
                     breaks = seq(1,12,1)) +
  labs(color = "Year")
```
- due to inconsisent unit, we use transaction to estimate the sales performance
- Dec has a huge number of transcations
- an obvious drop exists in 2016
```{r}
ggplot(trans %>% group_by(month, year) %>% 
         summarise(trans = sum(transactions),
                   shop_nr = length(unique(store_nbr)), # remove effect due to newly opened or closed stores
                   open_day = length(unique(date)), # remove effect due to inconsistance number of days in each month
                   trans_per_shop = trans/shop_nr/open_day) %>% 
         ungroup) +
  geom_line(aes(x=as.numeric(month), y=trans_per_shop, color = as.factor(year))) +
  ggtitle("Average number of daily transactions per store in each month") +
  scale_y_continuous(name = "Avg daily transactions per store") +
  scale_x_continuous(name = "Month",
                     breaks = seq(1,12,1)) +
  labs(color = "Year")
```
```{r}
ggplot(trans %>% group_by(wday) %>% 
         summarise(trans = sum(transactions),
                   shop_nr = length(unique(store_nbr)), # remove effect due to newly opened or closed stores
                   open_day = length(unique(date)), # remove effect due to inconsistance number of days
                   trans_per_shop = trans/shop_nr/open_day) %>% 
         ungroup) +
  geom_bar(aes(x=as.factor(wday), y=trans_per_shop), stat = "identity") +
  geom_text(aes(x = wday, y = 1, label = paste0("(", open_day,")",sep="")),
            hjust= 0.5, vjust= -2, size = 4, colour = 'black',
            fontface = 'bold') +
  ggtitle("Average number of daily transactions per store in each weeday") +
  scale_y_continuous(name = "Avg daily transactions per store") +
  scale_x_discrete(name = "Weekday (Opened days)", label = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))

```
It doesn't seem to be a high correlation between oil price and the number of transaction
```{r}
ggplot(trans %>% group_by(date) %>% 
         summarise(trans = sum(transactions),
                   store_opened = length(unique(store_nbr)),
                   avg_store_trans = trans / store_opened) %>%
         mutate(month = month(date),
                year = year(date)) %>%
         left_join(oil, by="date")) +
  geom_point(aes(x=dcoilwtico, y=avg_store_trans, color= as.factor(year)))+
  facet_wrap(~month)
```
Some stores have significant greater amount of transaction.
```{r}
ggplot(trans) +
  geom_smooth(aes(x=date, y=transactions, color=as.factor(store_nbr)), 
              method = "loess", span = 1/2, se = FALSE) +
  ggtitle("Total sales per store")
```

